### 2022-01-29 docs/ms.tex
gonna eventually send this to Sage.

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
\author{true}
\date{}

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: November 08, 2021; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }

@@

\usepackage{longtable}
\usepackage{booktabs}
\LTcapwidth=.95\textwidth
\linespread{1.05}
\usepackage{hyperref}
\usepackage{setspace}
\let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
\usepackage{colortbl}
\arrayrulecolor{black}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}
\newcolumntype{d}{S[input-symbols = ()]}

@@

% 
\begin{document}

@@

\noindent \small{This article introduces \texttt{\{peacesciencer\}}, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g.~democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at \url{http://svmiller.com/peacesciencer}.}

@@

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs researchers how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students, in particular, face unique challenges associated with these developments. Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

\texttt{\{peacesciencer\}} is designed to address these problems. Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers. Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, state-year, and even state-day data from scratch. Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. \texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts. All are done with the maximum possible transparency. The project is available for public view of Github (\url{https://github.com/svmiller/peacesciencer/}). The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created and the \texttt{R} directory informs the user how each function works. The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language. A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992) and standard state-year analyses of civil conflict onset (e.g. Fearon \& Laitin, 2003). This feature concludes with a discussion of what else \texttt{\{peacesciencer\}} is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

@@

\texttt{\{peacesciencer\}} is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars (Bennett, Poast \& Stam, 2019). NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade (Bennett \& Stam, 2000). No matter, \texttt{\{peacesciencer\}} is motivated by the following ideals that led to its creation.

For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.~dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. \texttt{\{peacesciencer\}} is primarily born from this question about how to create dyad-year and state-year data from scratch. The underlying code that produces these data types is available online and \texttt{\{peacesciencer\}} converts these lines of code into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of \texttt{\{peacesciencer\}}'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

Third, the creation of the data and the analysis of the data are increasingly becoming one continuous process. Not too long ago, it used to be the case that researchers had to download a data set, or create one from scratch (possibly in a spreadsheet or through a program like EUGene). After downloading or constructing the data, the researcher then opened a specialty program for statistical analysis (e.g.~SAS, SPSS, Stata) to recode raw data into a form suitable for analysis before running a statistical model that regresses some outcome on a set of covariates. Current research practices still resemble this process, but the steps between them are no longer as large as they were in the past. Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program. \texttt{\{peacesciencer\}}, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and Rstudio (its free-for-use integrated desktop environment {[}IDE{]}).

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f. Bowers \& Voors, 2016). This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a), scrape \emph{any} type of data from the internet (Boehmke, 2016), and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b). \texttt{\{peacesciencer\}} embraces this. This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Fifth, statistical software should be kind to graduate students in peace science with the aforementioned points in mind. Graduate students just beginning their career in peace science scholarship are tasked to learn the substantive material, the characteristics of the underlying data in wide use in the community, and the quantitative methods and statistical software available to them, effectively all at the same time. A student interested in civil conflict scholarship will need to learn the UCDP-PRIO armed conflict data in some detail at the same time they learn about the theoretical arguments that informed the creation of the data, likely at the same time they take quantitative methods courses that show them how to load and explore data in a chosen statistical software. At the same time that creating data is increasingly blurred with coding and processing data, students are simultaneously learning to explore data for themselves as they read about the analyses of the data. \texttt{\{peacesciencer\}} aims to make this process easier for graduate students using the features of the R programming language. Graduate students can use \texttt{\{peacesciencer\}} to learn about peace science data and computer programming in R at the same time.

Finally, the creation and presentation of data in peace science should be 100\% robust and transparent, which \texttt{\{peacesciencer\}} takes seriously in the following ways. The website for \texttt{\{peacesciencer\}} has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. \texttt{\{peacesciencer\}} subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge). The project's Github contains a publicly available \texttt{data-raw} directory that shows how every data set included (and processed) in \texttt{\{peacesciencer\}} was created. The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way. Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions. \texttt{\{peacesciencer\}} takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

@@

\texttt{\{peacesciencer\}} comes with a fully developed suite of built-in functions for generating the most widespread forms of peace science data. Each function uses raw or pre-processed data included in \texttt{\{peacesciencer\}}. For example, \texttt{create\_dyadyears()} transforms the raw state system membership data for either the Correlates of War or Gleditsch-Ward system into the full universe of post-1816 dyad-years. \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler, Miller \& Little (2016), which, as the \texttt{data-raw} directory on the project's Github shows, is itself derived from the dispute-level and participant-level versions of the data. \texttt{\{peacesciencer\}} also has some data innovations. It has an original data set on capital-to-capital distances, calculating for distance in kilometers between state capitals using the ``Vincenty'' method (i.e.~``as the crow flies'') that also accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018). It also features an innovation of democracy data, providing reasonable estimates of democracy using the Marquez (2016) method of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010).\footnote{\texttt{\{peacesciencer\}} comes with a vignette that explains just how much unnecessary missingess pervades data on democracy and how this might adversely affect inferences of inter-state or intra-state conflict.} As of writing, \texttt{\{peacesciencer\}} comes with the following functionality described in Table \ref{tab:tab-functionality}.

\caption{\label{tab:tab-functionality}Features of \{peacesciencer\}}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{>{}l|>{}l|>{}c|>{}c}
\hline
\textbf{Data} & \textbf{Description} & \textbf{Type} & \textbf{Function}\\
\hline
(CoW, G-W) Dyad-Year Data & Create dyad-years from state system membership data &  & \ttfamily{create\_dyadyears()}\\
\arrayrulecolor{white}
\hline
(CoW, G-W) State-Year Data & Create state-years from state system membership data &  & \ttfamily{create\_stateyears()}\\
\arrayrulecolor{white}
\hline
(CoW, G-W) State-Day Data & Create state-days from state system membership data &  & \ttfamily{create\_statedays()}\\
\arrayrulecolor{white}
\hline
Archigos (v. 4.1) & Add Archigos political leader information & D, S & \ttfamily{add\_archigos()}\\
\arrayrulecolor{white}
\hline
ATOP Alliances (v. 5.0) & Add ATOP alliance data & D & \ttfamily{add\_atop\_alliance()}\\
\arrayrulecolor{white}
\hline
Capital-to-Capital Distances & Add capital-to-capital distance & D, S & \ttfamily{add\_capital\_distance()}\\
\arrayrulecolor{white}
\hline
Convert CoW codes to G-W codes & Add matching CoW codes for G-W states by year & D, S & \ttfamily{add\_ccode\_to\_gw()}\\
\arrayrulecolor{white}
\hline
Convert G-W codes to CoW codes & Add matching G-W codes for CoW states by year & D, S & \ttfamily{add\_gwcode\_to\_cow()}\\
\arrayrulecolor{white}
\hline
CoW National Capabilities (v. 5.0) & Adds capabilities of states by year & D, S & \ttfamily{add\_nmc()}\\
\arrayrulecolor{white}
\hline
CoW Direct Contiguity (v. 3.2) & Adds contiguity information by year & D, S & \ttfamily{add\_contiguity()}\\
\arrayrulecolor{white}
\hline
CoW Alliances (v. 4.1) & Adds CoW alliance data by year & D & \ttfamily{add\_cow\_alliance()}\\
\arrayrulecolor{white}
\hline
CoW IGOs (v. 3.0) & Adds CoW IGO data by year & D, S & \ttfamily{add\_igos()}\\
\arrayrulecolor{white}
\hline
CoW Major Powers (v. 2016) & Adds CoW major power by year & D, S & \ttfamily{add\_cow\_majors()}\\
\arrayrulecolor{white}
\hline
CoW MIDs (v. 5.0) & Add CoW MID information to data & D & \ttfamily{add\_cow\_mids()}\\
\arrayrulecolor{white}
\hline
CoW Trade Data (v. 4.0) & Add CoW trade information to data & D, S & \ttfamily{add\_cow\_trade()}\\
\arrayrulecolor{white}
\hline
CoW Wars & Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data & D & \ttfamily{add\_cow\_wars()}\\
\arrayrulecolor{white}
\hline
CREG (v. 1.02) & Add ethnic/religious fractionalization/polarization by year & D, S & \ttfamily{add\_creg\_fractionalization()}\\
\arrayrulecolor{white}
\hline
Democracy & Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions & D,S & \ttfamily{add\_democracy()}\\
\arrayrulecolor{white}
\hline
GML MIDs (v. 2.1.1) & Add Gibler-Miller-Little MIDs to data & D & \ttfamily{add\_gml\_mids()}\\
\arrayrulecolor{white}
\hline
Minimum Distance (v. 2016) & Adds minimum distance data by year & D, S & \ttfamily{add\_minimum\_distance()}\\
\arrayrulecolor{white}
\hline
Peace Years & Calculate peace years for various disputes & D, S & \ttfamily{add\_peace\_years()}\\
\arrayrulecolor{white}
\hline
Rugged/Mountainous Terrain & Add 'ruggedness'/'mountainous' terrain data by year & D, S & \ttfamily{add\_rugged\_terrain()}\\
\arrayrulecolor{white}
\hline
(Surplus/Gross) Domestic Product & Add Anders et al. (2020) GDP/SDP simulations by year & D, S & \ttfamily{add\_sdp\_gdp()}\\
\arrayrulecolor{white}
\hline
Strategic Rivalries & Add Thompson and Dreyer (2012) strategic rivalry data by year & D, S & \ttfamily{add\_strategic\_rivalries()}\\
\arrayrulecolor{white}
\hline
UCDP Armed Conflicts (v. 20.1) & Add UCDP armed conflict data by year & S & \ttfamily{add\_ucdp\_acd()}\\
\arrayrulecolor{black}
\hline
\end{tabular}
\end{table}

\texttt{\{peacesciencer\}}'s coverage focuses mostly on data that are released as standalone data sets for download, especially those in the Correlates of War or Gleditsch-Ward ecosystem of data. Data that can be obtained from a stable advanced programming interface---like the World Bank, for example---can be obtained through those other means (e.g. Arel-Bundock, 2021a). Its coverage will assuredly expand with new additions of interest to the peace science community.

>>>>

  \setmonofont[]{Fira Code}

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}}
\author{true}
\date{}

@@

\title{\texttt{\{peacesciencer\}}: An R Package for Quantitative Peace Science Research\thanks{Replication files are available on the author's Github account (\url{https://github.com/svmiller/peacesciencer}). \textbf{Current version}: January 29, 2022; \textbf{Corresponding author}: \href{mailto:steven.v.miller@gmail.com}{\nolinkurl{steven.v.miller@gmail.com}}.}  }

@@

\usepackage{longtable}
\usepackage{booktabs}
\usepackage{csquotes}
\LTcapwidth=.95\textwidth
\linespread{1.05}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}
\newcolumntype{d}{S[input-symbols = ()]}

@@

% 
% Optional code chunks ------
% SOURCE: https://stackoverflow.com/questions/50702942/does-rmarkdown-allow-captions-and-references-for-code-chunks

\begin{document}

@@

\noindent \small{This article introduces \texttt{\{peacesciencer\}}, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular \texttt{\{tidvyerse\}} packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g.~democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at \url{http://svmiller.com/peacesciencer}.}

@@

This manuscript tackles a recurring problem for researchers in the peace science community. True research reproducibility is best achieved creating data from scratch, though no published guide exists that informs researchers how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. Researchers may additionally spend too much time reproducing old code for standard information that goes into any dyadic or monadic analysis---like contiguity relationships and democracy---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students and other beginners in the field face unique challenges associated with these developments. Students just learning peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

\texttt{\{peacesciencer\}} is designed to address these problems. Built around the free and open source R programming language, \texttt{\{peacesciencer\}} contains a suite of data and functions for creating data of interest to researchers. Researchers can use \texttt{\{peacesciencer\}} to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch. Afterward, they can add a variety of standard information (e.g.~contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. \texttt{\{peacesciencer\}} comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts. All are done with the maximum possible transparency. The project is available for public view on Github (\url{https://github.com/svmiller/peacesciencer/}). The \texttt{data-raw} directory on the project's Github contains information and comments about how every data set was created. The function manuals (\url{http://svmiller.com/peacesciencer/reference}) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, \texttt{\{peacesciencer\}} not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists. Afterward, it provides an overview of what is included in \texttt{\{peacesciencer\}} to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use \texttt{\{peacesciencer\}} in the R programming language. A more comprehensive tutorial follows, showing how \texttt{\{peacesciencer\}} already has a suite of data and functions that can allow for effective replications of a ``dangerous dyads'' type analysis (Bremer, 1992), standard state-year analyses of civil conflict onset (e.g. Fearon \& Laitin, 2003), and even leader-year analyses of inter-state dispute initiation (e.g. Horowitz \& Stam, 2014). This feature concludes with a comparison of \texttt{\{peacesciencer\}} with other alternatives and a discussion of how \texttt{\{peacesciencer\}} can inform more reasoned design decisions for researchers in peace science.

@@

\texttt{\{peacesciencer\}} is motivated by the following observations and ideals that led to its creation. For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g.~dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. \texttt{\{peacesciencer\}} is primarily born from this question about how to create these data from scratch. The underlying code that produces these data types is available online and \texttt{\{peacesciencer\}} converts these lines of code into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume a researcher finished a state-year analysis on the correlates of civil conflict onset a few years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of \texttt{\{peacesciencer\}}'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

Third, the creation of the data and the analysis of the data are increasingly becoming one continuous process. Not too long ago, it used to be the case that researchers had to download a data set, or create one from scratch (possibly in a spreadsheet or through a program like EUGene). After downloading or constructing the data, the researcher then opened a specialty program for statistical analysis (e.g.~SAS, SPSS, Stata) to recode raw data into a form suitable for analysis before running a statistical model that regresses some outcome on a set of covariates. Current research practices still resemble this process, but the steps between them are no longer as large as they were in the past. Software options exist that allow the researcher to load data, create data, clean data, analyze data, and present the results of the analysis all within one program. \texttt{\{peacesciencer\}}, by itself, does not do all these things, but it seamlessly connects the beginning of the research process to the end of the research process without needing to leave the increasingly popular R programming language and RStudio (its free-for-use integrated development environment {[}IDE{]}).

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer (c.f. Bowers \& Voors, 2016). This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that \texttt{\{peacesciencer\}} uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet (Arel-Bundock, 2021a) and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero (Arel-Bundock, 2021b). \texttt{\{peacesciencer\}} embraces this. This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

Finally, the creation and presentation of data in peace science should be 100\% robust and transparent, which \texttt{\{peacesciencer\}} takes seriously in the following ways. The website for \texttt{\{peacesciencer\}} has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. \texttt{\{peacesciencer\}} subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge). The project's Github contains a publicly available \texttt{data-raw} directory that shows how every data set included (and processed) in \texttt{\{peacesciencer\}} was created. The function manuals included \texttt{\{peacesciencer\}} contain ample documentation that clarify what each function is doing, what it returns to the user, and why it is doing it this way. Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions. \texttt{\{peacesciencer\}} takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

@@

\texttt{\{peacesciencer\}} comes with a fully developed suite of built-in functions for generating some of the most widespread forms of peace science data and populating the data with important variables that recur in many quantitative analyses. The core functionality of \texttt{\{peacesciencer\}} reduces to two broad categories of functions. These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis.\footnote{In both cases, SQL-like ``joins''---especially the left (outer) join---are the primary workhorse in these functions. A vignette on the package website, included in the appendix, explains these various joins, why they are used, and what issues may arise from them.} Table \ref{tab:tab-createfunctions} and Table \ref{tab:tab-addfunctions} list these core functions as of version 1.0, the version of this package slated for release alongside the manuscript.\footnote{\texttt{\{peacesciencer\}} also has miscellaneous ``helper'' functions that are not belabored in this manuscript. These include functions that allow the user to use \texttt{\{peacesciencer\}} functions with data created outside the package (\texttt{declare\_attributes()}), download external data for quicker use of other functions (\texttt{download\_extdata()}), and functions that communicate suggested citations (\texttt{ps\_cite()}), among a few other functions. See the package's website for more information (\url{http://svmiller.com/peacesciencer}).}

Table \ref{tab:tab-createfunctions} lists functions that create base data frames for a researcher starting an original project. They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user. For example, \texttt{create\_stateyears()} will generate the full universe of state-years from the Correlates of War (Correlates of War, 2011: v. 2016) or Gleditsch-Ward (Gleditsch \& Ward, 1999: v. 2017) system, encompassing all state-years to the most recently concluded calendar year, depending on the arguments supplied to the user in the function. \texttt{create\_leaderyears()} will generate the full universe of leader-years from the Archigos leader data (Goemans, Gleditsch \& Chiozza, 2009: v. 4.1), optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data. As of version 1.0, \texttt{\{peacesciencer\}} is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.\footnote{Leader-dyad-year data are too time-consuming to calculate each time and too size-prohibitive to include in the package. These data can be downloaded through \texttt{download\_extdata()} for much easier use.} A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g.~dyadic-dispute-year, leader-months, state-quarters).

\begin{longtable}[t]{>{}ll}
\caption{\label{tab:tab-createfunctions}Functions that Create Base Data}\\
\toprule
\textbf{Function} & \textbf{Description}\\
\midrule
\ttfamily{create\_dyadyears()} & Create dyad-years from state system membership data\\
\ttfamily{create\_leaderdays()} & Create leader-day data from Archigos\\
\ttfamily{create\_leaderdyadyears()} & Create leader-dyad-years data from Archigos\\
\ttfamily{create\_leaderyears()} & Create leader-years data from Archigos\\
\ttfamily{create\_statedays()} & Create state-days from state system membership data\\
\ttfamily{create\_stateyears()} & Create state-years from state system membership data\\
\bottomrule
\end{longtable}

Table \ref{tab:tab-addfunctions} lists the main functions in \texttt{\{peacesciencer\}} that add information or subset the number of rows of the data to just those of interest to the user, describing these functions and listing whether they are applicable to dyad-year (\texttt{D}), leader-year (\texttt{L}), leader-dyad-year (\texttt{LD}), state-year (\texttt{S}) data or specialty functions applicable to just the dyadic conflict data (\texttt{C}).\footnote{The ``whittle'' (\texttt{wc\_}) class of functions is a suite of functions that whittle dyadic dispute-years to true dyad-year data. They are effectively used in \texttt{add\_cow\_mids()} and \texttt{add\_gml\_mids()} for dyad-year data, but are offered here to allow users to employ their own case exclusion rules to create their own dyad-year conflict data. A vignette on \texttt{\{peacesciencer\}}'s website explains these case exclusion rules in some detail.} All these functions use raw or pre-processed data included in the package. For example, \texttt{add\_gml\_mids()} uses a dyadic dispute-year version of the MID data offered by Gibler et al. (2016: v. 2.2.1) and merges in information about whether there was an ongoing MID or MID onset in a dyad-year, leader-year, leader-dyad-year, or state-year. \texttt{\{peacesciencer\}} also has some data innovations included in these functions. For example, \texttt{add\_capital\_distance()} calculates distance between state capitals in kilometers using the Vincenty method (i.e.~``as the crow flies'') based on an original data set of state capitals that accounts for instances when capitals moved (e.g.~Brazil in 1960, Burundi in 2018). \texttt{add\_democracy()} does more than just add Polity data to a data set. The data underpinning the function feature an innovation in democracy data, providing reasonable estimates of democracy using the Marquez (2016) method of extending the Unified Democracy Scores (UDS) data (Pemstein, Meserve \& Melton, 2010) in addition to Polity estimates (Marshall, Gurr \& Jaggers, 2017: v. 2017) and V-Dem estimates (Coppedge et al., 2020: v. 10).\footnote{\texttt{\{peacesciencer\}} comes with a vignette that explains just how much unnecessary missingness pervades data on democracy in the absence of an innovation like this, and how this missingness might adversely affect inferences of inter-state or intra-state conflict.}

\begingroup\fontsize{9}{11}\selectfont
